---
title: "CAQ Posit Connect Cloud demonstrator"
subtitle: "Day 1 deployment probe"
author: "Cancer Alliance Queensland"
date: today
format:
  html:
    theme: flatly
    toc: true
    toc-location: left
    embed-resources: false
    code-fold: true
    fig-width: 8
    fig-height: 4.5
execute:
  echo: false
  warning: false
  message: false
---

<!--
=============================================================================
facility-report.qmd — CAQ Posit Connect Cloud demonstrator

STATUS: Day 1 deployment probe. Placeholder content.

Per D-024 this file keeps its name for the life of the project, so the Connect
Cloud deployment record created today is reused by every later push.

What this probe tests (I-001):
  3. Working directory at render time
  4. Sourcing shared code from R/
  5. Reading committed data from data/
  6. ggplot2 rendering on the build host

Note on fonts (D-022): no showtext, no font_add_google(). The report uses the
Quarto/Bootstrap default font stack, so it carries no render-time font
dependency. Only the Shiny app exercises the network font path.
=============================================================================
-->

```{r}
#| label: setup

# Explicit library() calls so renv detects dependencies (D-015). renv scans .qmd
# chunks as well as .R files, so this block contributes to the lockfile.
library(dplyr)
library(readr)
library(tibble)
library(ggplot2)
library(scales)
library(gt)

# Shared layer — same relative path used by app.R. This is the D-017 flat-root
# assumption under test.
source("R/smoke_shared.R")

probe_data <- smoke_read_data()
```

::: {.callout-important title="Synthetic data"}
All data in this report is synthetic and generated for demonstration purposes.
It does not describe real patients, real facilities, or real clinical
performance. Indicator definitions are plausible but are not endorsed (D-002).
:::

## Purpose of this document

This is a deployment probe, not a report. Its only job is to confirm that a
Quarto document and a Shiny application can both be published to Posit Connect
Cloud from a single repository sharing one `renv.lock`, and that both resolve
shared code and committed data identically after deployment.

The facility-level indicator content replaces everything below from Day 3.

## Environment diagnostics

Compare this table against the equivalent table in the deployed Shiny
application. **The working directory row is the one that matters** — if the two
products disagree, the shared `R/` layer cannot be shared and issue I-001 falls
back to a two-repository split.

```{r}
#| label: diagnostics

smoke_diagnostics(product = "Quarto report") |>
  gt::gt() |>
  gt::cols_label(check = "Check", value = "Value") |>
  gt::tab_options(table.font.size = gt::px(13))
```

## Data read check

Confirms that a CSV committed to `data/` is readable after deployment
(I-001, check 5).

```{r}
#| label: data-check

probe_data |>
  gt::gt() |>
  gt::fmt_percent(columns = rate, decimals = 1) |>
  gt::cols_label(
    facility_id = "Facility",
    year        = "Year",
    n_cases     = "Cases",
    n_events    = "Events",
    rate        = "Rate"
  ) |>
  gt::tab_options(table.font.size = gt::px(13))
```

## Graphics device check

Confirms `ggplot2` renders on the Connect Cloud build host (I-001, check 6).

```{r}
#| label: plot-check

smoke_plot(probe_data)
```

## Next steps

| Day | Milestone |
|---|---|
| 1 | This probe deployed; I-001 checklist completed |
| 2 | Synthetic data generator; shared prep, metrics and suppression layer |
| 3 | Facility report content replaces this document |
| 4 | Shiny dashboard content replaces the probe app |
| 5 | Persona switcher, polish, dress rehearsal |
