---
title: "Cancer Indicator Report"
subtitle: "Facility and Health Service performance, 2015--2024"
author: "Cancer Alliance Queensland — Reporting & Analysis"
date: today
format:
  html:
    toc: true
    toc-depth: 2
    toc-location: left
    theme: flatly
    fig-width: 8
    fig-height: 5
    code-fold: true
    code-summary: "Show the code"
    embed-resources: true
execute:
  echo: true
  warning: false
  message: false
---

```{r}
#| label: setup
#| echo: false

# --- Resolve the repository root (I-016) -------------------------------------
# The Shiny app runs from /cloud/project; this report renders in an ephemeral
# copy at /tmp/<name>/rendered-output-<id>. Different absolute paths, same tree
# structure. Rather than assume a fixed relative depth, walk up until the
# shared layer is visible. Correct under Connect Cloud, under a local render,
# and under RStudio's Render button.
.root <- base::normalizePath(".")
while (!base::file.exists(base::file.path(.root, "R", "metrics.R")) &&
       base::dirname(.root) != .root) {
  .root <- base::dirname(.root)
}
knitr::opts_knit$set(root.dir = .root)
```

```{r}
#| label: libraries
#| echo: false

# Explicit library() calls, never pacman::p_load() (D-015) — renv and rsconnect
# discover dependencies by static analysis and cannot see packages passed as
# bare symbols.
library(dplyr)
library(ggplot2)
library(gt)
library(plotly)
library(htmltools)

# The shared layer. Neither product calculates its own numbers (D-010, I-011).
# Quarto does not get Shiny's automatic R/ sourcing, so these are explicit
# (D-025).
source(base::file.path(.root, "R", "theme.R"))
source(base::file.path(.root, "R", "data_prep.R"))
source(base::file.path(.root, "R", "metrics.R"))
source(base::file.path(.root, "R", "suppression.R"))
source(base::file.path(.root, "R", "plots.R"))

caq <- caq_load_data(base::file.path(.root, "data"))
pat <- caq$patients
ind <- caq$indicators

SYNTHETIC_NOTE <- "All data in this report is synthetic. It describes no real patients, facilities or clinical performance."
```

::: {.callout-important}
## This report contains no real data

Every figure below is generated by a seeded script from invented data. It describes
**no real patients, no real facilities and no real clinical performance**. Facility and
Health Service names are fictional. Indicator definitions are plausible but are not
endorsed and carry no clinical validity.

The subject of this document is the publishing workflow, not the findings.
:::

## Overview

```{r}
#| label: overview
#| echo: false

tibble::tibble(
  Measure = c("Patients", "Diagnosis years", "Tumour streams",
              "Hospital and Health Services", "Treating facilities",
              "Patients who had surgery"),
  Value = c(
    base::format(base::nrow(pat), big.mark = ","),
    base::paste0(base::min(pat$diagnosis_year), "–", base::max(pat$diagnosis_year)),
    base::paste(base::levels(pat$tumour_stream), collapse = ", "),
    base::as.character(base::nrow(caq$hhs)),
    base::as.character(base::nrow(caq$facilities)),
    base::paste0(base::format(base::sum(pat$had_surgery), big.mark = ","),
                 " (", scales::percent(base::mean(pat$had_surgery), accuracy = 0.1), ")")
  )
) |>
  gt::gt() |>
  gt::tab_options(table.font.size = gt::pct(90)) |>
  gt::cols_align("left")
```

Results are presented as two five-year periods, **2015--19** and **2020--24**. The period
is derived once in the shared layer so that this report and the accompanying dashboard
can never disagree on where the cut falls.

## Indicators at a glance

All six indicators, statewide, by period.

```{r}
#| label: summary-table

summary_rows <- base::lapply(ind$indicator_id[ind$measure_type != "distribution"],
  function(id) {
    def <- caq_indicator(ind, id)
    res <- caq_calculate_indicator(pat, ind, id, group_by = "diagnosis_period")
    res |>
      dplyr::transmute(
        Indicator = def$label,
        Period    = base::as.character(diagnosis_period),
        Value     = caq_format_value(value, def$format),
        Patients  = base::format(denominator, big.mark = ",")
      )
  })

dplyr::bind_rows(summary_rows) |>
  tidyr::pivot_wider(names_from = Period, values_from = c(Value, Patients),
                     names_glue = "{Period} {.value}") |>
  gt::gt(rowname_col = "Indicator") |>
  gt::tab_spanner(label = "2015–19",
                  columns = c("2015-19 Value", "2015-19 Patients")) |>
  gt::tab_spanner(label = "2020–24",
                  columns = c("2020-24 Value", "2020-24 Patients")) |>
  gt::cols_label(
    "2015-19 Value"    = "Result",
    "2015-19 Patients" = "Patients",
    "2020-24 Value"    = "Result",
    "2020-24 Patients" = "Patients"
  ) |>
  gt::tab_source_note(SYNTHETIC_NOTE) |>
  gt::tab_options(table.font.size = gt::pct(90))
```

## Surgery rate by Health Service of residence

```{r}
#| label: ind01-hhs

def01 <- caq_indicator(ind, "IND-01")

surg_hhs <- caq_calculate_indicator(
  pat |> dplyr::filter(tumour_stream == "Lung"),
  ind, "IND-01", group_by = "hhs_residence_name"
)

caq_plot_indicator_bar(
  surg_hhs, "hhs_residence_name", def01,
  title    = "Lung: surgery rate by Health Service of residence",
  subtitle = "All diagnosed patients. Kurrajong and Melaleuca residents highlighted.",
  caption  = SYNTHETIC_NOTE,
  highlight = c("Kurrajong", "Melaleuca")
)
```

::: {.callout-note}
## Why this indicator has no facility breakdown

A patient is attached to a treating facility **only if they had surgery**. Grouping the
surgery rate by facility would therefore silently drop everyone who did not have surgery
from the denominator — the very patients the indicator is about.

This is enforced in metadata rather than by convention: `indicator_definitions.csv`
declares `facility_dims_allowed`, the calculation layer refuses the grouping, and the
dashboard never offers it. Surgery rate and stage at diagnosis are analysable by Health
Service of residence only.
:::

## Stage at diagnosis

```{r}
#| label: ind06-stage

stage_hhs <- caq_calculate_indicator(
  pat |> dplyr::filter(tumour_stream == "Lung"),
  ind, "IND-06", group_by = "hhs_residence_name"
)

caq_plot_distribution(
  stage_hhs, "hhs_residence_name", "stage",
  title    = "Lung: stage at diagnosis by Health Service of residence",
  subtitle = "\"Unknown\" is shown as a category and is counted in the denominator.",
  caption  = SYNTHETIC_NOTE
) + ggplot2::coord_flip()
```

Stage is recorded as **Unknown** for a small share of patients. Unknown is presented as a
category and included in every denominator, rather than being dropped — removing a
category from a distribution while still presenting it as summing to 100% would
misrepresent the population the bar describes.

```{r}
#| label: stage-completeness
#| echo: false

caq_calculate_indicator(pat, ind, "IND-06", group_by = "tumour_stream") |>
  dplyr::filter(stage == "Unknown") |>
  dplyr::transmute(
    `Tumour stream` = base::as.character(tumour_stream),
    `Stage unknown` = scales::percent(value, accuracy = 0.1),
    `Patients`      = base::format(denominator, big.mark = ",")
  ) |>
  gt::gt() |>
  gt::tab_header(title = "Stage completeness") |>
  gt::tab_options(table.font.size = gt::pct(90))
```

## Facility-level indicators

Surgical-cohort indicators, by facility and period. Cells with fewer than five patients
are suppressed.

```{r}
#| label: facility-table

def04 <- caq_indicator(ind, "IND-04")

fac_tbl <- caq_calculate_indicator(
  pat, ind, "IND-04", group_by = c("facility_name", "diagnosis_period")
) |>
  caq_suppress(mode = "table")

fac_caption <- caq_suppression_caption(fac_tbl, mode = "table")

fac_tbl |>
  dplyr::transmute(
    Facility = base::as.character(facility_name),
    Period   = base::as.character(diagnosis_period),
    Rate     = value,
    Patients = denominator
  ) |>
  tidyr::pivot_wider(names_from = Period, values_from = c(Rate, Patients),
                     names_glue = "{Period} {.value}") |>
  gt::gt(rowname_col = "Facility") |>
  gt::fmt_percent(columns = dplyr::contains("Rate"), decimals = 1) |>
  gt::fmt_number(columns = dplyr::contains("Patients"), decimals = 0, use_seps = TRUE) |>
  gt::sub_missing(missing_text = "—") |>
  gt::tab_header(title = def04$label) |>
  gt::tab_source_note(SYNTHETIC_NOTE) |>
  gt::tab_source_note(if (base::is.null(fac_caption)) "" else fac_caption) |>
  gt::tab_options(table.font.size = gt::pct(85))
```

## Funnel plot: variation between facilities

```{r}
#| label: funnel

def03 <- caq_indicator(ind, "IND-03")

funnel_res <- caq_calculate_indicator(pat, ind, "IND-03", group_by = "facility_name") |>
  caq_suppress(mode = "funnel", label_col = "facility_name")

funnel_note <- caq_suppression_caption(funnel_res, mode = "funnel")

p_funnel <- caq_plot_funnel(
  funnel_res, "facility_name", def03,
  title    = "Hospital-acquired complications by facility",
  subtitle = "Pooled rate with 95% and 99.8% control limits"
)

# plotly for funnel plots only (D-014) — points are unidentifiable without
# tooltips. Note ggplotly() drops subtitle and caption, so the footnote is
# rendered separately below (I-004).
plotly::ggplotly(p_funnel, tooltip = c("text", "x", "y"))
```

```{r}
#| label: funnel-note
#| echo: false

# The footnote states what suppression ACTUALLY did, rather than describing the
# rule as though it had fired. At facility level the smallest denominator is in
# the hundreds, so nothing falls below the threshold here (D-012).
htmltools::tags$p(
  style = "font-size: 0.85em; color: #555; margin-top: -0.5em;",
  "Solid line: pooled rate across all facilities. Dashed lines: 95% control limits. ",
  "Dotted lines: 99.8% control limits. ",
  if (!base::is.null(funnel_note)) {
    base::paste(funnel_note, "")
  } else {
    base::paste0(
      "No facility falls below the five-patient suppression threshold at this level ",
      "of aggregation — the smallest facility contributes ",
      base::format(base::min(funnel_res$denominator), big.mark = ","),
      " surgical patients across the ten years. "
    )
  },
  SYNTHETIC_NOTE
)
```

Control limits widen at small denominators because a rate measured on few patients is
less precise, not because those facilities are more variable in any meaningful sense. A
point inside the limits is consistent with the pooled rate.

## Prolonged length of stay: a closer look at one facility

Ironbark Southbank's colorectal prolonged-stay rate, compared with the all-facility rate,
across the two reporting periods.

```{r}
#| label: s1-two-periods

s1 <- caq_calculate_indicator(
  pat |> dplyr::filter(tumour_stream == "Colorectal", facility_name == "Ironbark Southbank"),
  ind, "IND-04", group_by = "diagnosis_period"
)

overall <- caq_calculate_indicator(
  pat |> dplyr::filter(tumour_stream == "Colorectal"),
  ind, "IND-04", group_by = "diagnosis_period"
)

dplyr::bind_rows(
  s1 |> dplyr::mutate(Group = "Ironbark Southbank"),
  overall |> dplyr::mutate(Group = "All facilities")
) |>
  dplyr::transmute(
    Group,
    Period   = base::as.character(diagnosis_period),
    Rate     = value,
    Patients = denominator
  ) |>
  tidyr::pivot_wider(names_from = Period, values_from = c(Rate, Patients),
                     names_glue = "{Period} {.value}") |>
  gt::gt(rowname_col = "Group") |>
  gt::fmt_percent(columns = dplyr::contains("Rate"), decimals = 1) |>
  gt::fmt_number(columns = dplyr::contains("Patients"), decimals = 0, use_seps = TRUE) |>
  gt::tab_header(title = "Prolonged length of stay, colorectal surgery") |>
  gt::tab_source_note(SYNTHETIC_NOTE) |>
  gt::tab_options(table.font.size = gt::pct(90))
```

The period comparison shows a deterioration at this facility against a stable
all-facility rate. What it cannot show is the **shape** of that change: when it began,
whether it is still occurring, or whether it has since been addressed. A five-year
average answers none of those questions, and a reader of this report has no way to tell
the difference between a steady decline and a spike that has already resolved.

That is a limitation of static period reporting, not of the underlying data. The
accompanying dashboard presents the same figures year by year.

## Notes on method

```{r}
#| label: definitions
#| echo: false

ind |>
  dplyr::transmute(
    Indicator  = label,
    Cohort     = base::ifelse(cohort == "surgical", "Surgical patients", "All diagnosed"),
    `By facility?` = base::ifelse(facility_dims_allowed, "Yes", "No"),
    Definition = definition_text
  ) |>
  gt::gt() |>
  gt::tab_options(table.font.size = gt::pct(85)) |>
  gt::cols_width(Definition ~ gt::pct(45))
```

**Suppression.** Cells describing fewer than five patients are suppressed. Aggregation
runs across all facilities first and suppression is applied afterwards, so pooled rates
and control limits are computed from the complete data. In tables the count is withheld
and the row retained; in funnel plots the point is retained and the facility name
withheld.

```{r}
#| label: suppression-status
#| echo: false
#| output: asis

# State plainly whether the rule actually fired anywhere in this document. A
# report that describes a disclosure control it never applied invites the
# reader to assume it was applied silently.
report_min <- base::min(fac_tbl$denominator, funnel_res$denominator, na.rm = TRUE)

base::cat(base::sprintf(
  paste0("In this report no cell falls below that threshold: the smallest group presented ",
         "contains %s patients. The five-year periods and facility totals used here ",
         "aggregate well above the level at which suppression becomes necessary — ",
         "the aggregation is itself acting as disclosure control. The accompanying ",
         "dashboard, which presents figures by individual year and tumour stream, does ",
         "reach cells below the threshold and suppresses them.\n"),
  base::format(report_min, big.mark = ",")
))
```

**Eligibility.** Indicators restricted to surgical patients count only those patients in
the denominator. Patients who did not have surgery are not attached to a facility and are
excluded from facility-level measures by definition, not by filtering.

**Reproducibility.** The data is generated by a seeded script committed alongside this
report. The package environment is pinned locally with `renv`, and those pins are carried
into `manifest.json`, which is what the publishing platform provisions from.

::: {.callout-important}
`r SYNTHETIC_NOTE` No conclusion in this document should be drawn about any real service.
:::
